Supplementary information: In-silico molecular enrichment and clearance of the human intracranial space

### 2025-06-05 paper/supplementary.tex
Updates from Overleaf
Also note about vasomotion antegrade in supplementary.

\begin{abstract}
  Supplementary methods and discussion accomanying the main text of
  \emph{In-silico molecular enrichment and clearance of the human
  intracranial space}.
\end{abstract}

@@

\flushbottom
\maketitle

@@

\input{supplementary_main.tex}

>>>>

\begin{abstract}
  Supplementary methods and discussion accompanying the main text of
  \emph{In-silico molecular enrichment and clearance of the human
  intracranial space}.
\end{abstract}

@@

\linenumbers
\flushbottom
\maketitle

@@

\section{Supplementary methods}

These sections provide more detailed descriptions of the mathematical models and numerical approximations considered.

\subsection{Mixed-dimensional transport and flow equations}
\label{sec:app:more_maths}

\subsubsection{Notation}

In terms of geometrical domains, we consider the parenchyma
$\Omega_{\rm PAR} \subset \R^3$ and CSF spaces $\Omega_{\rm CSF}
\subset \R^3$ with $\Omega = \Omega_{\rm PAR} \cup \Omega_{\rm CSF}$
(Fig.~1A, Fig.~2A--C). The coordinates in these
3D domains is denoted by $x$. The interface between the parenchyma and
CSF spaces is given by $\partial \Omega_{\rm PAR} \cap \partial
\Omega_{\rm CSF}$ and we separate it as two parts: the surface of the
lateral ventricles $\Gamma_{\rm LV}$ and the remaining pial interface
$\Gamma_{\rm pia}$ (thus $\partial \Omega_{\rm CSF} \bigcap \partial
\Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm pia}$). The
remaining, outer, boundary of $\Omega_{\rm CSF}$ is again separated
into two parts: $\Gamma_{\rm SSAS}$ represents the lower interface
towards the spinal subarachnoid space (SSAS), while $\Gamma_{\rm AM}$
is the outer interface towards the arachnoid and dura
membranes. $\Gamma_{\rm AM}$ is further subdivided into its lower and
upper parts: $\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$. The boundary
towards the spinal cord is denoted by $\Gamma_{\rm{SC}}$ and given by
$\Gamma_{\rm{SC}} = \partial \Omega_{\rm PAR} \setminus \partial
\Omega_{\rm CSF}$.

In addition, we consider two sets of perivascular networks: a
periarterial network $\Lambda_a$ represented by the (connected)
centerlines $\Lambda^i_{a}$ of the arterial tree, and a perivenous
network $\Lambda_v$ associated with the centerlines $\Lambda^i_{v}$
of the veins. Recall that we represent the vascular domains as the
union of cylindrical vessels of radius $R_1^i$ surrounding the
centerlines $\Lambda^i$. Moreover, we consider the surrounding
perivascular spaces as the union of annular cylinders $\Omega^i$ of
inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
width $R_2^i - R_1^i$, and remark that we interpret $R^i_2$ and $R^i_1$ as temporal averages (fixed in time), as the pulsations considered in \Cref{sec:app:peristalsis} are beyond the temporal resolution of our model. We denote the outer lateral surface of the
periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
respectively. We omit the subscript $a$ or $v$ when referring to any
such network, vessel segment or perivascular outer surface. We assume
that $\Lambda$ is parametrized by the coordinate $s$, and with a minor
abuse of notation, simply write $s \in \Lambda$ to represent the point
${\bm \lambda}(s)$ on $\Lambda$ corresponding to $s$. Finally, we view
each perivascular network both as a geometric domain and as a directed
graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
connections as nodes $\mathcal{V}$.

\subsubsection{Solute transport and exchange in intracranial domains (3D) and perivascular networks (1D)}
\label{sec:app:transport_eqs}

In this section, we describe the mathematical model (eq.~(1)), the
associated definitions, interface and boundary conditions in further
detail. We refer to the main text for explicit parameter values while
giving the general, abstract form here, and to the
reference~\cite{masri2024modelling, laurino2019derivation} for the
derivation and analysis of this 3D-1D model. Recall that we model a
concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
each of these domains, $c$ satisfies
%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here. We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
\begin{equation}
  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
  %= f
   = 0
  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
  \label{eq:3d_pde}
\end{equation}
In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$). In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}. Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$. $\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivascular networks $\Lambda_a, \Lambda_v$. To summarize,
\begin{equation*}
\phi =  \begin{cases}
  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
  \end{cases}, \; 
D = \begin{cases}
  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
  \quad \bm u  = \begin{cases}
  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
\end{cases}, \; 
\xi = \begin{cases}
  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
\end{cases} . 
\end{equation*}
In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp. $|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp. $\Omega_{\rm{CSF}}$). If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.

Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
\begin{equation*}
  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
\end{equation*}
where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$. Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the term $\delta_\Gamma$ is concentrated on the outer lateral surfaces of the PVSs, and defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$
%and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$
such that
\begin{equation*}
  \langle \xi (\overline{c} - \hat c ) \delta_\Gamma, v \rangle =
  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
\end{equation*}
under the assumption that $\xi$ is constant on each $\partial
\Theta_2^i$.
%% , since:
%% \begin{multline}
%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
%% \end{multline}
%% \mer{MER: The previous was:
%% \begin{equation}
%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
%% \end{equation}
%% }

On the interface between the brain and the CSF spaces
$\Gamma_{\mathrm{pia}} \cup \Gamma_{\rm LV}$, we prescribe the
following interface conditions, which represent a semi--permeable
interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and
$c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
%% \begin{subequations}
%% \begin{alignat}{2}
%% (-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} &= - (- D (\phi \nabla c_{\rm{CSF}}) + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n} &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
%% (-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
%% \end{alignat} 
%% \end{subequations}
\begin{subequations}
  \begin{alignat}{2}
  (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
  &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
  (- D \nabla (\phi  c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
\end{alignat} 
  \label{eq:3d_interface}%
\end{subequations}%
where $\bm{n} = \bm{n}_{\rm PAR}$ is the normal vector field defined
over the interface, outward-pointing when viewed from $\Omega_{\rm
  PAR}$, and $\beta_{\mathrm{pia}} \geqslant 0$ is a permeability
constant.

We supplement~\eqref{eq:3d_pde} and~\eqref{eq:3d_interface} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
\begin{subequations}
\begin{alignat}{2}
(D \nabla (\phi c) - \bm u c ) \cdot \bm{n} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \Gamma_{\rm SSAS},  \\ 
% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\Gamma_{\rm SAS}} &= 0 &&  \quad  \mathrm{on} \quad \Gamma_{\rm SAS}^{\mathrm{out}}  \\ 
(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta_{\rm exit} c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \cup \Gamma_{\rm SC}  .
\end{alignat} 
\label{eq:3d_bcs}
\end{subequations}
where again $\bm n$ denotes the outward-pointing boundary normal.

The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
space averaged over each perivascular cross-section and is governed by 
\begin{equation}
  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s \hat c  ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
\label{eq:1d_pde}
\end{equation}
where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
associated PVS $\Omega^i$, the area of the cross-section
$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
and $\hat{u}$ is a convective velocity representing an (average) CSF
flow velocity in the axial direction of the perivascular spaces.

To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
\hat{c}|_{\Lambda^i}$, and for $y\in\mathcal{B}$ let $\mathcal{E}(y)$ denote the set of
(two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
impose continuity of the concentrations, we set that the
concentrations when viewed from each edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}^j(y)
  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
  \qquad \foralls y \in \mathcal{B} .
  \label{eq:conc_continuity}
\end{equation}
Moreover, to ensure mass conservation, we set that the flux
going in and out at nodes should add to zero, or more precisely, that
for each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{\Lambda^i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n^i(y)
  = 0.
  \label{eq:flux_conservation}
\end{equation}
Here the (normal ``vector") function $n^i$ takes the values in $\left\{-1, 1\right\}$ and 
defines an orientation of vertices of edge $\Lambda^i$. Specifically, for an edge $\Lambda^i$ oriented from $y_{\rm in}$ to $y_{\rm out}$ we set 
\begin{equation}
  n^i(y_{\rm in}) = 1,\quad
  n^i(y_{\rm{out}}) = -1. %\quad
  %\mk{n^i(y) = 0 \quad \foralls y \in 
  %\mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
\label{eq:def:sign}
\end{equation}
%\mk{Why do we need the normal on non-IO vertices?}

Finally, we impose a homogeneous Neumann condition at all network end
nodes to augment~\eqref{eq:1d_pde}:
\begin{equation}
    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{N}.
    \label{eq:1d_bcs}
\end{equation}
This no-flux condition states that the network end nodes represent
barriers for molecular efflux into or out of the perivascular network,
and thus that all solute exchange between the PVSs and their
surroundings takes place via the lateral outer PVS surface and is
regulated by the exchange parameter $\xi$ cf.~\eqref{eq:3d_pde}
and~\eqref{eq:1d_pde}. Larger particles have been observed to
accumulate within the PVS as the surface arteries penetrate into the
brain parenchyma~\cite{mestre2018flow, bedussi2018paravascular},
and~\eqref{eq:1d_bcs} is appropriate to represent such
behavior. However, to represent a continuously extending PVSs also
along penetrating vessels, this condition should be revisited.

\subsubsection{Modelling CSF flow via the incompressible Stokes equations}
\label{sec:app:csf_flow}
The time-independent, incompressible Stokes equations model the flow
of an incompressible Newtonian fluid at low Reynolds numbers, and read
as follows: over the domain $\Omega_{\rm CSF} \subset \R^3$, find the
velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
the pressure field $p : \Omega_{\rm CSF} \rightarrow \R$ such that
\begin{subequations}
  \begin{alignat}{2}
    - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:momentum_equation} \\ 
    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF}
    \label{eq:divergence_equation} .
  \end{alignat}
  \label{eq:stokes}%
\end{subequations}%
In addition, we impose the following boundary conditions to model flow
induced by CSF production in the choroid plexus with $\Gamma_{\rm AM-U}$ as the main efflux site:
\begin{subequations}
  \begin{alignat}{2}
    ( \mu \nabla \bm u \cdot \bm{n} - p \bm n) \cdot \bm n  &= -R_0 ( \bm u \cdot \bm n ), 
    \quad \bm u \cdot \bm t = 0
    && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
    \bm u &= 0 
    && \text{ on } \Gamma_{\mathrm{AM-L}}\cup\Gamma_{\rm{pia}}\cup\Gamma_{\rm{SSAS}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, 
    \quad \quad \bm u \cdot \bm t = 0 \quad 
    && \text{ on } \Gamma_{\rm{LV}} ,  
  \end{alignat}
  \label{eq:stokes_bcs}%
\end{subequations}%
where $\bm n$ and $\bm t$ denote the unit outward normal and tangent vectors to the boundary respectively,  $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$. We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:app:dispersion}.

%% We consider non-zero convection in the CSF space $u_{\rm CSF} \not = 0$ in \eqref{eq:multi_transport_3d} resulting from steady state CSF production.  Recall that $\Omega_{\mathrm{CSF}}$ represents the CSF space and that $\Gamma_{\mathrm{skull^{u/l}}}$ and $\Gamma_{\mathrm{pia}}$, $\Gamma_{\mathrm{LV}}$ represent the boundaries facing the upper and lower parts of the dura, the pia and the lateral ventricles, respectively.  Assuming CSF production in the choroid plexus located at the lateral ventricle wall, we solve the  steady state Stokes equations for the velocity and the pressure $(\bm u_{\mathrm{CSF}}, p_{\mathrm{CSF}})$ in the CSF domain:

\subsubsection{Steady flow in perivascular networks induced by pressure differences}
\label{sec:app:hydraulic_network}

To model steady flow induced by pressure differences between end nodes
in a perivascular network $\Lambda$, with edges $\{\Lambda^i\}_i$,
internal nodes $\mathcal{B}$ and end nodes $\mathcal{N}$, we consider
the following system of hydraulic network
equations~\cite{tithof2022network, daversin2022geometrically,
  gjerde2024directional}. The unknowns are the PVS flux $\hat{q} :
\Lambda \rightarrow \R$ and pressure $\hat{p} : \Lambda
\rightarrow \R$, which represent the fluid flux and the average
pressure across cross-sections of the PVSs, respectively. These are
defined for each $\Lambda^i$ by solving:
\begin{subequations}
\begin{alignat}{2}
  \hat{q} + \frac{A \kappa}{\mu} \partial_{s} \hat{p} & = 0 &&
  \quad \text{ in } \Lambda^i , \\ 
  - \partial_s \hat{q} & = 0 &&
  \quad \text{ in } \Lambda^i ,  
\end{alignat} 
\label{eq:1d_flow}%
\end{subequations}%
where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}: 
\begin{equation}
\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
\end{equation}
The PVS flux $\hat{q}$ and the associated average PVS velocity in the
axial direction $\hat{u}$ are directly related by the cross-sectional area
\begin{equation}
  \hat{q}^i = A^i \hat{u}^i \qquad \foralls \Lambda^i.
\end{equation}

In addition, to complete~\eqref{eq:1d_flow}, we impose continuity of
the fluid pressure and conservation of flux at each internal node. Write
$\hat{p}^i = \hat{p}|_{\Lambda^i}$. These conditions then read as: for
each $y \in \mathcal{B}$ with connected edges $\mathcal{E}(y)$:
\begin{subequations}
\begin{align}
  \hat{p}^i(y) &= \hat{p}^j(y) \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y), \\
  \sum_{\Lambda^i \in \mathcal{E}(y)} \hat{q}^i n^i(y) &= 0  .
\end{align}
\label{eq:1d_flow:bifurcations}
\end{subequations}
where $n^i$ is as defined by~\eqref{eq:def:sign}. Finally, to drive this flow, we impose given fluid pressures $p_0$ at the
network end nodes:
\begin{equation}
  \hat{p}(y) = p_0(y) \qquad y \in \mathcal{N} .
  \label{eq:1d_flow:bcs}
\end{equation}
In the case where modelling flow in the perivascular networks
$\Lambda_a, \Lambda_v$ induced by CSF production, we impose the fluid
pressure $p_{\rm CSF}$ computed in the CSF spaces $\Omega_{\rm CSF}$
(and its harmonic extensions into $\Omega_{\rm PAR}$) as the given
fluid pressures $p_0$. Note that we here consider steady PVS flow in a
non-moving domain; the effect of domain motion is addressed
in~\Cref{sec:app:peristalsis}.

\subsection{Numerical solution of the transport and flow equations}
\label{sec:app:details_numerical_method}

We solve the models presented in~\Cref{sec:app:more_maths} numerically
using finite element methods. The following subsections provide more
details, model-by-model.

\subsubsection{Computational mesh entities: definitions and common notation}

We consider a mesh $\mathcal{T} = \{ E \}$ of $\Omega = \Omega_{\rm
  PAR} \cup \Omega_{\rm CSF}$, consisting of tetrahedral mesh cells
$E$, and conforming to the domains $\Omega_{\rm PAR}$ and $\Omega_{\rm
  CSF}$ and to the CSF-brain interface $\Gamma_{\rm pia} \cup
\Gamma_{\rm LV}$. We denote the restriction of $\mathcal{T}$ to
$\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ by $\mathcal{T}_{\rm PAR}$
and $\mathcal{T}_{\rm CSF}$, respectively. The collection of all
interior facets (i.e.~triangular faces of the tetrahedral mesh cells)
in $\mathcal{T}_{\rm PAR}$ and $\mathcal{T}_{\rm CSF}$ are denoted by
$\mathcal{F}_{i, \rm PAR}$ and $\mathcal{F}_{i, \rm CSF}$,
respectively. We define the union of facets interior both to
$\mathcal{T}_{ \rm PAR}$ and $\mathcal{T}_{ \rm CSF}$ as
$\mathcal{F}_i = \mathcal{F}_{i, \rm PAR} \cup \mathcal{F}_{i, \rm
  CSF}$.

%Note in particular that $\mathcal{F}_i \cap \mathcal{F}_{pia}} =
%  \emptyset$.

%The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is
%denoted by $\mathcal{E}_{h}$.

\subsubsection{Finite element solution of coupled 3D-1D solute transport equations }

We consider the system of coupled 3D-1D solute transport equations
given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
conditions~\eqref{eq:3d_interface}, the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
these equations using an implicit finite difference scheme in time, 
a discontinuous Galerkin (DG) finite element method with upwinding in space for the 3D domain to accurately capture sharp boundary layers, and a continuous Galerkin method for the 1D networks. We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm, and accounting for the increased diffusivity due to dispersion).

We first consider the discretization of~\eqref{eq:3d_pde} and
introduce the discrete space
\begin{equation}
  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
\end{equation}
where $L^2(\Omega)$ is the space of square-integrable functions
defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
polynomials of total degree $\leqslant 1$ defined over the tetrahedra
$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
symmetric weighted interior penalty DG formulation, referring
to~\cite{ern2009discontinuous} and \cite[Section
  4.5.2.3]{di2011mathematical} for details on this
method. Recall that $\phi$ is constant in each domain
  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
Define for $c, v \in V_h^k$:
\begin{equation}
  a_h (c, v)
  =
  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right) 
   + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
  .
  \label{eq:dg_sip}
\end{equation}
In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
are defined as:
\begin{equation}
  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the
porosity-weighted diffusion coefficient given by
\begin{equation}
  \gamma_{D,F} = \frac{2\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
\end{equation}
and $\eta$ is a user-specified penalty parameter (we set $\eta=1000$). To discretize the convection term in~\eqref{eq:3d_pde}, we use
upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
references therein. For $c, v \in V_h^k$, define
\begin{equation}
  a_h^{\mathrm{up}}(c, v)
  = - \sum_{E \in \mathcal{T}} \int_E c\bm{u} \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
\end{equation}
Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
\begin{equation} \label{eq:3d_dg_disc}
  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
  = 
  %\int_{\Omega} f^n v +
  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
  %, \quad \forall v \in V_h^k
  .
\end{equation}

To discretize~\eqref{eq:1d_pde} with the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:1d_bcs}, we use the space of continuous piecewise
linear polynomials defined over $\Lambda$:
\begin{equation}
  \hat V_h
  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i) \}, 
\end{equation}
where $\mathrm{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$. The discrete formulation then reads: for $n = 1,2,\ldots, $ find $\hat c_h^n \in \hat V_h$ such that for all $\hat v \in \hat V_h^k$: 
\begin{align}\label{eq:1d_dg_disc}
\int_{\Lambda}  \frac{1}{\tau}(A  \hat c_h^n - A \hat c_h^{n-1}) \hat v  + \int_{\Lambda} \hat D A \partial_s \hat c_h^n  \partial_s \hat v^n - \int_{\Lambda} A \hat u \hat c_h^n \partial_s \hat v   +  \int_{\Lambda} \xi P (\hat c_h^n - \overline{c_h^n}) \hat v +a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) = 0.  
\end{align}
 In the above, if $\hat u$ is non--zero on $\Lambda^i$, then the artificial diffusion $a_h^{\mathrm{stab}}$ stabilization term is nonzero  and is given below, for more details see for  \cite[Section 12.6]{quarteroni2009numerical}. 
 \begin{align}
    a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) =  \frac{1}{\|\hat u\|_{L^2(\Lambda^i)}} 
 h_{\Lambda^i} \int_{\Lambda^i} A (\hat u \partial_s \hat c_h  ) (\hat u \partial_s \hat v),  
 \end{align}
 where $h_{\Lambda^i}$ is the mesh--size of $\Lambda^i$.
Note that the condition \eqref{eq:flux_conservation} is enforced weakly in the above formulation.

 \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} complemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:

 For $n = 1,2, \ldots $, find  $c_h^n \in V_h$ and $\hat c^n_h \in \hat V_h$ such that the coupled equations \eqref{eq:3d_dg_disc} and \eqref{eq:1d_dg_disc} hold for all $v \in V_h$ and for all $\hat v \in \hat V_h$. 
\subsubsection{Finite element solution of the incompressible Stokes equations}
